\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{enumitem}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    citecolor=blue!60!black,
    urlcolor=blue!60!black
}

\title{MAGELLAN: Autonomous Cross-Disciplinary Scientific Hypothesis Generation\\via Adversarial Multi-Agent Systems}

\author{
Alberto Trivero\\
Kakashi Venture Accelerator\\
Torino, Italy\\
\texttt{alberto@kakashi.ventures}\\
\url{https://magellan-discover.ai}
}

\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
We present MAGELLAN, an open-source multi-agent system that autonomously generates cross-disciplinary scientific hypotheses by connecting fields that lack mutual citation. Building on Swanson's ``Undiscovered Public Knowledge'' principle (1986), MAGELLAN replaces bibliometric methods with frontier large language models that have absorbed the literature of multiple disciplines. The system comprises 15 specialized agents: a Scout that autonomously identifies promising cross-field bridges using 10 formalized strategies, a Generator that creates hypotheses from parametric knowledge, and an adversarial pipeline (Critic with 9 attack vectors, Quality Gate with ${\sim}60\%$ rejection rate, independent validation by GPT-5.4~Pro and Gemini~3.1~Pro) that systematically eliminates weak candidates.

Over 22 sessions, MAGELLAN generated 273 hypotheses, of which 102 (37\%) survived the full adversarial pipeline. We computationally verify two representative hypotheses on public data: (1)~applying extreme value theory to the Meltome Atlas (48{,}000 proteins, 13 species)---the first such application, revealing universal Weibull-domain behavior but failing to confirm the predicted correlation with growth temperature due to measurement censoring; (2)~deriving the Cram\'{e}r-Rao bound for plant gravitropic sensing from published statolith parameters---confirmed, showing plants operate 1--6$\times$ above the fundamental physical limit, the first application of statistical estimation theory to plant gravitropism. Meta-learning across sessions reveals that mathematical constraints transfer more reliably than analogies, and fully disjoint field pairs produce substantially better hypotheses (84\% vs.\ 30\% survival).

Unlike proprietary systems (Google AI Co-Scientist, FutureHouse), MAGELLAN is fully open-source (Apache~2.0), fully autonomous in target selection, and publishes all results as CC0 public domain. Code, hypotheses, methodology, and computational verifications are publicly available.
\end{abstract}

% ======================================================================
% SECTIONS 1-3: Introduction, Related Work, Architecture
% ======================================================================
\input{sections_1_to_3}

% ======================================================================
% SECTIONS 4-7: Evaluation, Case Studies, Discussion, Conclusion
% ======================================================================
\input{sections_4_to_7}

% ======================================================================
% REFERENCES
% ======================================================================
\bibliographystyle{plainnat}

\begin{thebibliography}{30}

\bibitem[Swanson(1986)]{swanson1986}
Swanson, D.R. (1986).
Undiscovered public knowledge.
\emph{The Library Quarterly}, 56(2), 103--118.

\bibitem[Swanson(1988)]{swanson1988}
Swanson, D.R. (1988).
Migraine and magnesium: eleven neglected connections.
\emph{Perspectives in Biology and Medicine}, 31(4), 526--557.

\bibitem[Tenopir et~al.(2009)]{tenopir2009}
Tenopir, C. et~al. (2009).
Electronic journals and changes in scholarly article seeking and reading patterns.
\emph{Aslib Proceedings}, 61(1), 5--32.

\bibitem[Rein et~al.(2023)]{rein2023}
Rein, D. et~al. (2023).
GPQA: A graduate-level Google-proof Q\&A benchmark.
\emph{arXiv preprint arXiv:2311.12022}.

\bibitem[Ginkgo Bioworks(2026)]{ginkgo2026}
Ginkgo Bioworks (2026).
Cell-free protein synthesis optimization via GPT-5 collaboration.
\emph{Ginkgo Technical Report}.

\bibitem[Gottweis et~al.(2025)]{gottweis2025}
Gottweis, J. et~al. (2025).
Towards an AI co-scientist.
\emph{Google DeepMind Technical Report}.

\bibitem[DeepMind(2025)]{deepmind2025aletheia}
DeepMind (2025).
Aletheia: Solving open mathematical problems with AI.
\emph{Google DeepMind Technical Report}.

\bibitem[Ji et~al.(2023)]{ji2023}
Ji, Z. et~al. (2023).
Survey of hallucination in natural language generation.
\emph{ACM Computing Surveys}, 55(12), 1--38.

\bibitem[OpenAI(2025)]{openai2025frontier}
OpenAI (2025).
FrontierScience-Research benchmark results.
\emph{OpenAI Technical Report}.

\bibitem[Si et~al.(2025a)]{si2025}
Si, C. et~al. (2025).
Can LLMs generate novel research ideas? A large-scale human study with 100+ NLP researchers.
\emph{ICLR 2025}.

\bibitem[Si \& Hashimoto(2025)]{si2025execution}
Si, C. \& Hashimoto, T. (2025).
The ideation-execution gap: Execution outcomes of LLM-generated versus human research ideas.
\emph{arXiv preprint arXiv:2506.20803}.

\bibitem[Si et~al.(2024)]{si2024}
Si, C. et~al. (2024).
Can LLMs generate diverse ideas? Measuring and improving idea diversity in LLM-generated scientific research ideas.
\emph{arXiv preprint}.

\bibitem[Smalheiser \& Swanson(2009)]{smalheiser2009}
Smalheiser, N.R. \& Swanson, D.R. (2009).
Arrowsmith: A computer tool for literature-based discovery.
In \emph{Bentham Science Publishers}.

\bibitem[Hristovski et~al.(2006)]{hristovski2006}
Hristovski, D. et~al. (2006).
Using literature-based discovery to identify disease candidate genes.
\emph{International Journal of Medical Informatics}, 75(12), 879--889.

\bibitem[Ghafarollahi \& Buehler(2024)]{ghafarollahi2024}
Ghafarollahi, A. \& Buehler, M.J. (2024).
SciAgents: Automating scientific discovery through multi-agent intelligent graph reasoning.
\emph{arXiv preprint arXiv:2409.05556}.

\bibitem[Mitchener et~al.(2025)]{mitchener2025}
Mitchener, L. et~al. (2025).
Kosmos: An AI scientist for autonomous discovery.
\emph{arXiv preprint arXiv:2511.02824}.

\bibitem[Lu et~al.(2024)]{lu2024}
Lu, C. et~al. (2024).
The AI Scientist: Towards fully automated open-ended scientific discovery.
\emph{ICLR 2025}.

\bibitem[Huang et~al.(2025)]{huang2025popper}
Huang, K., Jin, Y., Li, R., Li, M.Y., Cand\`{e}s, E. \& Leskovec, J. (2025).
Automated hypothesis validation with agentic sequential falsifications.
\emph{International Conference on Machine Learning (ICML 2025)}.
arXiv:2502.09858.

\bibitem[Swanson et~al.(2024)]{swanson2024virtuallab}
Swanson, K. et~al. (2024).
Virtual Lab: AI agents design new SARS-CoV-2 nanobodies with experimental validation.
\emph{arXiv preprint arXiv:2407.00510}.

\bibitem[Yang et~al.(2024b)]{yang2024moosechem}
Yang, B. et~al. (2024).
MOOSE-Chem: Large language models for rediscovering unseen chemistry scientific hypotheses.
\emph{ICLR 2025}.

\bibitem[Xiong et~al.(2025)]{xiong2025}
Xiong, R. et~al. (2025).
TruthHypo: Truthful hypothesis generation with LLMs.
\emph{IJCAI 2025}.

\bibitem[Jarz\k{a}b et~al.(2020)]{jarzab2020}
Jarz\k{a}b, A. et~al. (2020).
Meltome atlas---thermal proteome stability across the tree of life.
\emph{Nature Methods}, 17, 495--503.

\bibitem[B\'{e}rut et~al.(2018)]{berut2018}
B\'{e}rut, A. et~al. (2018).
Gravisensors in plant cells behave like an active granular liquid.
\emph{PNAS}, 115(20), 5123--5128.

\bibitem[Chauvet et~al.(2016)]{chauvet2016}
Chauvet, H. et~al. (2016).
Inclination not force is sensed by plants during shoot gravitropism.
\emph{Scientific Reports}, 6, 35431.

\bibitem[Berg \& Purcell(1977)]{berg1977}
Berg, H.C. \& Purcell, E.M. (1977).
Physics of chemoreception.
\emph{Biophysical Journal}, 20(2), 193--219.

\bibitem[Kawamoto \& Morita(2022)]{kawamoto2022}
Kawamoto, N. \& Morita, M.T. (2022).
Gravity sensing and responses in the coordination of the shoot gravitropic setpoint angle.
\emph{New Phytologist}, 236, 1445--1460.

\bibitem[Fisher \& Tippett(1928)]{fisher1928}
Fisher, R.A. \& Tippett, L.H.C. (1928).
Limiting forms of the frequency distribution of the largest or smallest member of a sample.
\emph{Mathematical Proceedings of the Cambridge Philosophical Society}, 24(2), 180--190.

\bibitem[Gnedenko(1943)]{gnedenko1943}
Gnedenko, B.V. (1943).
Sur la distribution limit\'{e}e du terme maximum d'une s\'{e}rie al\'{e}atoire.
\emph{Annals of Mathematics}, 44(3), 423--453.

\bibitem[Barato \& Seifert(2015)]{barato2015}
Barato, A.C. \& Seifert, U. (2015).
Thermodynamic uncertainty relation for biomolecular processes.
\emph{Physical Review Letters}, 114(15), 158101.

\bibitem[Grippin et~al.(2025)]{grippin2025}
Grippin, A.J., Marconi, C., Copling, S. et~al. (2025).
mRNA vaccines sensitize tumors to immune checkpoint blockade.
\emph{Nature}, 647(8089), 488--497.

\bibitem[DeCasien et~al.(2026)]{decasien2026}
DeCasien, A.R., Aronoff, J.E., Mallott, E.K. et~al. (2026).
Gut microbiota from large-brained primates upregulate brain oxidative phosphorylation genes.
\emph{Proceedings of the National Academy of Sciences}, 123(2), e2426232122.

\bibitem[Cosgrove et~al.(2025)]{cosgrove2025}
Cosgrove, B.D., Bounds, L.R., Taylor, C.K. et~al. (2025).
Mechanoenhancers: genomic regulatory elements activated by extracellular matrix stiffness.
\emph{Science}, 390(6778), eadl1988.

\bibitem[DeepMind(2025b)]{deepmind2025alphaevolve}
DeepMind (2025).
AlphaEvolve: A Gemini-powered coding agent for designing advanced algorithms.
\emph{Google DeepMind Technical Report}.

\bibitem[Schmidgall et~al.(2025)]{schmidgall2025}
Schmidgall, S. et~al. (2025).
Agent Laboratory: Using LLM agents as research assistants.
\emph{arXiv preprint arXiv:2501.04227}.

\bibitem[Li et~al.(2025)]{li2025survey}
Li, J. et~al. (2025).
A survey of AI scientists: Toward a six-stage scientific research paradigm.
\emph{arXiv preprint arXiv:2510.23045}.

\bibitem[Phan et~al.(2025)]{phan2025hle}
Phan, L. et~al. (2025).
Humanity's Last Exam.
\emph{Nature}.

\bibitem[Majumder et~al.(2024)]{majumder2024discoverybench}
Majumder, B.P. et~al. (2024).
DiscoveryBench: Towards data-driven discovery with large language models.
\emph{NeurIPS 2024}.

\bibitem[Liu et~al.(2025)]{liu2025researchbench}
Liu, J. et~al. (2025).
ResearchBench: Benchmarking LLMs in scientific discovery via inspiration-driven research.
\emph{arXiv preprint arXiv:2503.21370}.

\end{thebibliography}

\end{document}
